---
title: "Warmup: data wrangling with the Mets"
format: pdf
editor: source
---

**Group members:**

**Instructions:** Work with a neighbor on the following activity. I will collect the handout at the end of class, and it will be part of your class participation grade. You will be graded only on effort -- it is ok if you don't finish all the questions, or get them all correct.

# Baseball data

The `Teams` dataset in contains 3045 rows and 48 columns, and describes yearly statistics for baseball teams between 1871 and 2023. 

Here are the first few rows and columns of the dataset:

```{r, echo=F, message=F, warning=F}
library(Lahman)
library(tidyverse)

head(Teams[,1:13])
```

## Questions

1. The following chunk of code is run, producing the output below. Explain what the `select` function is doing.

```{r, eval=F}
Teams |>
  select(yearID, franchID, W, L)
```

```{r, echo=F}
Teams |>
  select(yearID, franchID, W, L) |>
  head()
```

\vspace{3cm}

2. Now suppose we want to get *only* the data for the New York Mets (franchise ID "NYM"), between 1998 and 2018. Fill in the code; the first few rows of the desired output are shown below.

```{r, eval=F}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(...)
```

```{r, echo=F}
Teams |>
  select(yearID, franchID, W, L) |>
  filter(franchID == "NYM",
         yearID >= 1998, yearID <= 2018) |>
  head()
```
