---
title: "Warmup: summary statistics"
format: pdf
editor: source
---

\vspace{-2.5cm}

**Group members:**

**Instructions:** Work with a neighbor on the following activity. I will collect the handout at the end of class, and it will be part of your class participation grade. You will be graded only on effort -- it is ok if you don't finish all the questions, or get them all correct.

# Calculating summary statistics

The `diamonds` dataset from the `ggplot2` package contains information on a variety of characteristics of different diamonds (cut, color, clarity, price, etc.). Here are the first few rows:

```{r, echo=F, message=F}
library(tidyverse)
head(diamonds) |>
  as.data.frame()
```

1. Write code using `dplyr` functions to calculate the mean and standard deviation of the `carat`, `depth`, and `price` columns, producing the output shown below:

```{r, echo=F}
diamonds |>
  summarize(mean_carat = mean(carat),
            sd_carat = sd(carat),
            mean_depth = mean(depth),
            sd_depth = sd(depth),
            mean_price = mean(price),
            sd_price = sd(price))
```





