---
title: "Activity: Practice pivoting with the dog data"
format: html
editor: source
---

**Instructions:** 

* Work with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta279-f25.github.io/class_activities/ca_05_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

# Dog data

On the first day of class, we discussed an experiment which investigated whether interacting with dogs can help exam stress for college students. After cleaning the raw measurements, the experiment data looks like this:

```{r setup, include=F}
library(tidyverse)

raw_data <- read.csv("https://raw.githubusercontent.com/ciaran-evans/dog-data-analysis/main/dog_data_raw.csv")

cleaned_data <- raw_data %>%
  mutate(across(starts_with("PN"), ~ replace(.x, .x > 5, NA)),
         across(starts_with("L2"), ~ replace(.x, .x > 4, NA)),
         across(starts_with("SC"), ~ replace(.x, .x > 6, NA)),
         across(paste("L1_", c(1, 5, 6, 9, 10, 15, 16, 19, 20), sep=""),
                function(x){5-x}),
         across(paste("L2_", c(1, 5, 6, 9, 10, 15, 16, 19, 20), sep=""),
                function(x){5-x}),
         across(paste("SC1_", c(3, 6, 7, 9, 11, 13, 15, 17, 18, 20), sep=""),
                function(x){7-x}),
         across(paste("SC2_", c(3, 6, 7, 9, 11, 13, 15, 17, 18, 20), sep=""),
                function(x){7-x}),
         Engagement1 = 6 - HO1_2,
         Engagement2 = 6 - HO2_2)


cleaned_data <- cleaned_data %>%
  mutate(pa_pre = rowMeans(select(cleaned_data,
                                     num_range("PN1_", range = c(3,5,7,8,10))),
                              na.rm = TRUE),
         pa_post = rowMeans(select(cleaned_data,
                                     num_range("PN2_", range = c(3,5,7,8,10))),
                              na.rm = TRUE), 
         happiness_pre = rowMeans(select(cleaned_data, num_range("HA1_", range = c(1,2,3))), na.rm = TRUE),
         happiness_post = rowMeans(select(cleaned_data, num_range("HA2_", range = c(1,2,3))), na.rm = TRUE),
         sc_pre = rowMeans(select(cleaned_data, starts_with("SC1_")), na.rm = TRUE),
         sc_post = rowMeans(select(cleaned_data, starts_with("SC2_")), na.rm = TRUE),
         fs_pre = rowMeans(select(cleaned_data, starts_with("F1_")), na.rm = TRUE),
         fs_post = rowMeans(select(cleaned_data, starts_with("F2_")), na.rm = TRUE),
         stress_pre = S1_1,
         stress_post = S2_1,
         homesick_pre = HO1_1,
         homesick_post = HO2_1,
         lonely_pre = rowMeans(select(cleaned_data, starts_with("L1_")), na.rm = T),
         lonely_post = rowMeans(select(cleaned_data, starts_with("L2_")), na.rm = T),
         na_pre = rowMeans(select(cleaned_data,
                                     num_range("PN1_", range = c(1,2,4,6,9))),
                              na.rm = TRUE),
         na_post = rowMeans(select(cleaned_data,
                                     num_range("PN2_", range = c(1,2,4,6,9))),
                              na.rm = TRUE)) %>%
  dplyr::select(RID, GroupAssignment, pa_pre, 
                pa_post, happiness_pre, happiness_post, 
                sc_pre, sc_post, fs_pre, fs_post, stress_pre, stress_post,
                homesick_pre, homesick_post,
                lonely_pre, lonely_post, na_pre, na_post)

# write_csv(cleaned_data, "cleaned_dog_data.csv")
```

```{r}
head(cleaned_data)
```

Each row of the cleaned data represents one student, with columns for student ID (`RID`), their experimental group assignment (`Control` means handler-only contact, `Indirect` means a dog was present, `Direct` means direct contact with a dog), and their pre- and post-intervention scores for different wellbeing and illbeing measurements.

For example, `happiness_pre` is the student's assessed happiness before the intervention, and `happiness_post` is the student's assessed happiness after the intervention.

The data can be imported into R with the following code:


```{r, eval=F}
library(readr)

cleaned_data <- read_csv("https://sta279-f25.github.io/class_activities/cleaned_dog_data.csv")
```

## Pivoting

1. Reshape the `cleaned_data` in R so it looks like this (note that I am only displaying the first few rows):

```{r, echo=F}
cleaned_data |>
  pivot_longer(cols = -c(RID, GroupAssignment),
               names_to = c("measurement", "stage"),
               names_sep = "_",
               values_to = "score") |>
  head() |>
  as.data.frame()
```

2.

Reshape the `cleaned_data` in R so it looks like this (note that I am only displaying the first few rows):

```{r, echo=F}
cleaned_data |>
  pivot_longer(cols = -c(RID, GroupAssignment),
               names_to = c(".value", "stage"),
               names_sep = "_") |>
  head() |>
  as.data.frame()
```

3. Using your reshaped data from question 2, fit a linear model with `happiness` as the response variable and `GroupAssignment` and `stage` as the explanatory variables. Would you have been able to fit this model without pivoting the data?


