---
title: "Warmup: data wrangling across columns"
format: pdf
editor: source
---

\vspace{-2.5cm}

**Group members:**

**Instructions:** Work with a neighbor on the following activity. I will collect the handout at the end of class, and it will be part of your class participation grade. You will be graded only on effort -- it is ok if you don't finish all the questions, or get them all correct.


## starwars data

The `starwars` data (from the `dplyr` package) has information on characteristics (height, mass, birth year, etc.) for many different Star Wars characters. Here are the first few rows and columns of the data:

```{r echo=F, message=F}
library(tidyverse)

head(starwars) |>
  select(1:7) |>
  as.data.frame()
```

1. Without listing any column names explicitly, write code to calculate the median of each numeric variable in the `starwars` dataset.

\vspace{3cm}

2. Your friend also writes code for question 1, which returns the following output. Explain why your friend is getting `NA` for each summary statistic.

```{r, echo=F}
starwars |>
  summarize(across(where(is.numeric),
                   list("median" = median),
                   .names = "{.fn}_{.col}"))
```




