---
title: "Activity: practice with lists"
format: html
editor: source
---

## Practice with lists

1. Each of the following lists contains the vector `c(2, 7, 9)`. Determine the correct list indexing syntax to extract that vector from the list (your code should return just the vector, *not* a list containing the vector).

```{r}
x1 <- list(c(2, 7, 9))
x2 <- list(list(c(2, 7, 9)))
x3 <- list(c(2, 7, 9), list("a", "b"))
x4 <- list(c("a", "b"), list(list(c(2, 7, 9))))
```

**Solution:**

```{r}
x1[[1]]
x2[[1]][[1]]
x3[[1]]
x4[[2]][[1]][[1]]
```


2. Create a list `x5` such that the vector `c(2, 7, 9)` can be extracted with `x5[[3]][[2]][[1]]`.

**Solution:** Here is one example:

```{r}
x5 <- list("a", "b", list("c", list(c(2, 7, 9))))
x5[[3]][[2]][[1]]
```


3. Create a list `x` in R such that:

* `x[[1]]` returns the function `mean`
* `x[[2]]` returns the function `sd`
* `x[[3]][[1]]` returns the vector `c(0, 1, 2)`
* `x[[3]][[2]]` returns a function which calculates the cube root

**Solution:**

```{r}
x <- list(mean, sd, list(c(0, 1, 2), function(y) y^(1/3)))
```



