---
title: "Class activity solutions"
format: html
---

## Probability simulation: theater seating

Suppose there are exactly 100 seats in a movie theater. Further, suppose the movie theater has imposed a new ticketing system where each person’s ticket has an assigned seat number. Consider the case where the movie is sold-out (all 100 seats will be taken) and the first person in the theater lost their ticket and takes a random seat. Each subsequent person attempts to go to their seat. If their seat is open, they take it. Otherwise, to avoid confrontation, they choose a random seat. 

1. Conduct a simulation to assess the probability that the last person into the theater will get their assigned seat.

**Solution:**

For loop approach:

```{r}
set.seed(771)

n_people <- 100 # number of people in the theater
nsim <- 1000 # number of simulations to estimate probability
results <- rep(NA, nsim)
seats <- 1:n_people

for(i in 1:nsim){
  # vector to store which seats are taken
  # taken[i] is 0 when the seat is free
  taken <- rep(0, n_people)
  
  # first person randomly chooses a seat
  choice <- sample(seats, 1)
  taken[choice] <- 1
  
  # now go through everyone else (except the last person)
  for(j in 2:(n_people - 1)){
    
    # if the seat is free, take their seat. Otherwise,
    # randomly choose a seat from the ones available
    choice <- ifelse(taken[j] == 0, j, 
                     sample(seats[taken == 0], 1))
    taken[choice] <- 1
  }
  
  results[i] <- taken[n_people]
}

mean(results)
```

**Notes:**

* Make sure to use difference indices for the two `for` loops! Here I have used `i` and `j`
* In the inner for loop, I used `2:(n_people - 1)` to do `2, 3, ..., 99`. This is *different* from `2:n_people - 1`, which would give us `1, 2, ..., 99`.
* There are lots of other ways to code this question (e.g., you could use a counter to record the results, rather than a vector)
* It is crucial to treat the first person separately from the rest


Map approach: One run of the simulation is actually easier to do with the for loop, because the first person is treated differently than the others, and we need to update the `taken` vector. However, we can write a function to do one run of the simulation, then iterate it with `map`.

```{r, message=F}
library(tidyverse)
set.seed(771)

n_people <- 100 # number of people in the theater
nsim <- 1000 # number of simulations to estimate probability

take_seats <- function(n){
  seats <- 1:n
  taken <- rep(0, n)
  
  # first person randomly chooses a seat
  choice <- sample(seats, 1)
  taken[choice] <- 1
  
  # now go through everyone else (except the last person)
  for(j in 2:(n - 1)){
    
    # if the seat is free, take their seat. Otherwise,
    # randomly choose a seat from the ones available
    choice <- ifelse(taken[j] == 0, j, 
                     sample(seats[taken == 0], 1))
    taken[choice] <- 1
  }
  
  taken[n_people]
}

results <- map_dbl(1:nsim, function(i) take_seats(n_people))

mean(results)
```



