---
title: "Activity: Strings and regular expressions II"
format: html
editor: source
---

**Instructions:** 

* Work with a neighbor to answer the following questions 
* To get started, download the [class activity template](https://sta279-f25.github.io/class_activities/ca_17_template.qmd) file
* When you are finished, render the file as an HTML and submit the HTML to Canvas (let me know if you encounter any problems)

# Episode titles

The following code loads a data frame containing episode information from Season 11 of the British TV show *Taskmaster*:

```{r, message=F, warning=F}
library(tidyverse)
episode_info <- read_csv("https://sta279-f25.github.io/class_activities/taskmaster_episodes.csv")
```

The resulting data frame contains columns for the task, task description, episode, contestant, and score.

Examining the `episode` column, we see that the entries are strings that look like

```
"Episode 1: It's not your fault. (18 March 2021)"
```

That is, the string contains three different pieces of information: the episode number, the episode title, and the air date. In this activity, we will pull out each of these pieces of information.


## Questions

1. Extract just the episode numbers from the `episode` column.

2. Now we want to extract the episode title for each entry. Use positive lookaheads and lookbehinds to extract the episode titles from the `episode` column. *Hint:* Parentheses `( )` are special characters in regular expressions. To match a *literal* parenthesis, you will need to use escape characters -- that is, `\\(` and `\\)`

3. Finally, use positive lookaheads and lookbehinds to extract the episode air dates from the `episode` column.

4. Using your answers to the previous questions, modify the `episode_info` dataset so that the `episode` column is split into three different columns: episode number, episode title, and episode air date. Here is some example output:

```{r, echo=F}
episode_info |>
  mutate(episode_num = str_extract(episode, "\\d+"),
         title = str_extract(episode, "(?<=: ).+(?=\\()"),
         air_date = str_extract(episode, "(?<=\\().+(?=\\))")) |>
  select(episode_num, title, air_date) |>
  as.data.frame() |>
  head()
```

# Phone numbers

Below is a vector containing 10 phone numbers:

```{r}
phone_numbers <- c("(336) 703-2910",
                   "(336) 703-2665",
                   "(336) 703-2920",
                   "(336) 703-2930",
                   "(336) 703-2940",
                   "(336) 703-2950",
                   "(336) 703-2960",
                   "(336) 703-2970",
                   "(336) 703-2980",
                   "(336) 703-2990")
```

5. Use string functions and regular expressions to convert these phone numbers to the following format: `336-703-2910`

For question 5, try doing this with and without back references. Helpful string functions for this question include `str_remove_all`, `str_replace`, and `str_replace_all`




